---
title: "Chapters 7 and 8 Squamate Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(geiger)
require(phytools)
require(RCurl)
require(diversitree)
```

## Reading in the data files

First we read in the data files.

```{r}

sqTree<-read.tree(text=getURL("https://raw.githubusercontent.com/lukejharmon/pcm/master/datafiles/squamate.phy"))
plot(sqTree)

sqData<-read.csv(text=getURL("https://raw.githubusercontent.com/lukejharmon/pcm/master/datafiles/brandley_table.csv"))


```


## Simulate binary character on tree

This code generates plots like Figure 7.4

```{r}

qMatrix<-cbind(c(-1, 1), c(1, -1))*0.001
sh_slow<-sim.history(sqTree, qMatrix, anc="1")
plotSimmap(sh_slow, pts=F, ftype="off")
add.simmap.legend(leg=c("limbed", "limbless"), colors=c("black", "red"), x=0.024, y =23, prompt=F)


qMatrix<-cbind(c(-1, 1), c(1, -1))*0.01
sh_fast<-sim.history(sqTree, qMatrix, anc="1")
plotSimmap(sh_fast, pts=F, ftype="off")


qMatrix<-cbind(c(-0.02, 0.02), c(0.005, -0.005))
sh_asy<-sim.history(sqTree, qMatrix, anc="1")

plotSimmap(sh_asy, pts=F, ftype="off")


```

## Find the limbless species

Brandley et al.'s data has limb measurements. We will get our discrete character by counting species with zero-length fore- and hind limbs as limbless. This is different from the original analysis in Brandley et al., which counts things like spurs as "limbs" - and so our results might differ from theirs a bit.

```{r}

limbless<-as.numeric(sqData[,"FLL"]==0 & sqData[,"HLL"]==0)
sum(limbless)

# get names that match
nn<-sqData[,1]
nn2<-sub(" ", "_", nn)
names(limbless)<-nn2


```

### Fit Mk model

We can fit a symmetric Mk model to these data using both likelihood and MCMC

```{r}

# likelihood
td<-treedata(sqTree, limbless)
dModel<-fitDiscrete(td$phy, td$data)

# MCMC
mk_diversitree<-make.mk2(force.ultrametric(td$phy), td$data[,1])
simplemk<-constrain(mk_diversitree, q01~q10)
er_bayes<-mcmc(simplemk, x.init=0.1, nsteps=10000, w=0.01)
hist(er_bayes[-(1:1000),"q10"])
quantile(er_bayes[-(1:1000),"q10"], c(0.025, 0.975))
mean(er_bayes[-(1:1000),"q10"])
```



## Ancestral state reconstruction

We can do a basic ASR analysis.

```{r}

td<-treedata(sqTree, limbless)

limbless2<-as.numeric(td$data)
names(limbless2)<-rownames(td$data)

aa<-rerootingMethod(sqTree, limbless2, model="ER")

plot(sqTree,no.margin=TRUE,edge.width=2, show.tip.label=FALSE)
nodelabels(pie=aa$marginal.anc,piecol=palette()[1:4], cex=0.3)

mm<-match(sqTree$tip.label, names(limbless2))
tiplabels(pie=to.matrix(limbless2[mm],colnames(aa$marginal.anc)), piecol=palette()[1:4],cex=0.1)
add.simmap.legend(leg=c("limbed", "limbless"), colors=c("black", "red"), x=0.024, y =23, prompt=F)


```

## Compare ER and ASY models

Now we can compare the fit of a model where gains and losses occur at the same rate to one where the rates differ.

```{r}

# now we can try to compare the fit of some models using fitDiscrete
dModel2<-fitDiscrete(td$phy, td$data, model="ARD")

# are these different?
dModel$opt$aicc
dModel2$opt$aicc
# not really.

# Bayesian analysis
st2<-force.ultrametric(td$phy)
bayesFit<-mcmc(mk_diversitree, x.init=c(0.001610655, 0.003824740), lower=c(0.000001, 0.000001), upper=c(1,1), nsteps=10000, w=1)


plot.PostDist <- function(data, col=c("dodgerblue4", "#eaab00", "FF6C91"), pdf=FALSE, file=NA, ...){
	
	cc <- colnames(data)
	par <- vector()
	for (i in 1:length(cc)){
		tmp <- rep(cc[i], nrow(data))
		par <- c(par, tmp)
	}

	estimate <- vector()
	for (j in 1:length(cc)){
		tmp2 <- data[,cc[j]]
		estimate <- c(estimate, tmp2)
	}

	data.by.par <- cbind.data.frame(estimate, par)
	colnames(data.by.par) <- c("estimate", "par")
	
	col <- col[1:length(cc)]
	
	if (pdf == TRUE){
		pdf(file=file, ...)
		
		ggplot(data.by.par, aes(x = estimate), environment= environment()) + geom_density(alpha = 0.5, aes(fill=par)) + scale_fill_manual(values=col) + ggtitle("Posterior Distribution")
	
	}
	
	p <- ggplot(data.by.par, aes(x = estimate), environment= environment()) + geom_density(alpha = 0.5, aes(fill=par)) + scale_fill_manual(values=col) + ggtitle("Posterior Distribution")
	
	print(p)
	
	if (pdf == TRUE){
		dev.off()
	}

}
require(ggplot2)
plot.PostDist(bayesFit[,2:3])
```

```
